Foregrounds in Fourier Space for Redshifted 21~cm Observatories

### 2015-02-05 ms.tex
Complete section 5 on MWA data, modeling analysis
from a4e4df6:
============

Add figure8.eps, figure9.eps to git version control

In ms.tex,

(*) Make minor changes to captions of figures 5 and 7
(*) Write subsection on foreground signature analysis under S5
(*) Complete section 5 now consisting of the entire MWA data, modeling,
comparison and foreground analysis including the "pitchfork" section
(*) Remove section "Results"

\begin{figure*}[htb]
\centering
\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith. Antenna shapes used are: dipole (left), phased array (middle), and dish (right). Refer to text for details. The strength and occupancy of the power patterns are correlated with those of delay power spectra. The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish. The phased array has intermediate levels of contamination. Foreground power close to the horizon delay limits in all three cases are significant even on long baselines. The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales. The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
\label{fig:aperture-shapes}
\end{figure*}

@@

\begin{figure*}[htb]
\centering
\includegraphics[width=\linewidth]{figure7.eps}
\caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The modeling matches the data to a level consistent with the uncertainties in foreground models and the antenna beam. \label{fig:fhd-sim-comparison}}
\end{figure*}

@@

We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

@@

\section{Results}\label{sec:delay-spectrum-analysis}

@@

\item {\it Sky Model}: Our model of the sky consists of compact and diffuse emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}. It also consists of localized regions of strong emission such as the Galactic plane and Galactic center. % The resulting sky emission is anisotropic. 
In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission. We will investigate the signatures of each component and their potential impact on {\it EoR window} contamination.

@@

\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the tile power pattern tile changes in any observing mode that tracks the {\bf pointing center}. In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}. % We will show the secondary lobe structures in the tile power pattern are tied to the foreground signatures in Fourier space.

@@

\subsection{Diffuse Foreground Signatures}\label{sec:diffuse}

Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the {\bf delay power spectra} simulated with only the diffuse foreground model. The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \includegraphics[width=\linewidth]{f8.eps}
\includegraphics[width=\linewidth]{fig9.eps}
\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

Here we examine in detail some examples of interesting features observed with the diffuse foreground model.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.

\subsubsection{Diffuse Emission Delay Spectra}\label{sec:diffuse-features}

Figure~\ref{fig:noiseless-dsm-delay-spectrum} {\bf also shows diffuse emission in the primary field of view manifesting in the {\it off--zenith} pointing as a branch at $\tau>0$, and at $\tau=0$ in the {\it zenith} pointing. The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith}. As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission. Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.

@@

{\bf In typical radio interferometric measurements, due to the Fourier relation between sky brightness and measured visibilities, only short antenna spacings are sensitive to emission from large size scales such as the diffuse component in our study (angular scales $\gtrsim$~0.85\arcdeg). However, we note a very interesting signature of diffuse emission revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing, and at both horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m). This is a consequence of wide--field measurements, where long baselines are foreshortened due to projection effects towards off--axis emission, especially near the horizon. Consequently, they become sensitive to diffuse emission on large size scales. This holds for short baselines as well and is evident at all LSTs in our simulations}. Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines. It decreases in strength with increasing baseline length as expected but is present in all baseline orientations.

{\bf \citet{thy13} and \citet{pob13} have found a similar feature in their simulations and data respectively. In the latter, this feature is described as ``edge brightening'' and attributed to the Galactic plane near the horizon. From their discussion, it is unclear what fraction of power in that feature arises from ubiquitous diffuse emission. On the other hand, \citet{thy13} statistically modeled point sources below the confusion limit and attributed the feature to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission. This must be true for diffuse emission as well. Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis. Consequently, the increased size scales (or solid angles) containing higher levels of integrated emission, and spatial scales inversely proportional to the foreshortened baseline lengths become matched}.

\subsection{Compact Foreground Signatures}\label{sec:compact}

Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the {\bf delay power spectra} of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively. In contrast to the {\bf delay power spectra} of diffuse emission, compact emission manifests as a center--heavy structure in either pointing.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \includegraphics[width=\linewidth]{f9.eps}
\includegraphics[width=\linewidth]{fig10.eps}
\caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission. Compact emission gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}

The amplitude response of an interferometer to a point source is, to first order, flat with respect to baseline length. Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary lobe of the power pattern is centered at zenith.

@@

When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, {\bf both of which are significant}. This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum}, and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines. The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to 
uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \includegraphics[width=\linewidth]{f10.eps}
\includegraphics[width=\linewidth]{fig11.eps}
\caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}). The antenna spacing is specified in each panel along East, North and Upward directions. The solid red, cyan, and black lines denote contributions from diffuse, compact, and composite foreground models respectively. Vertical dotted lines mark the horizon delay limits. {\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure. The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum. This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations elsewhere}. \label{fig:pitchfork-baselines}}
\end{figure}

@@

Near the horizon delay limits at shorter antenna spacing, the diffuse component is brighter relative to the compact component. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay. In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude. We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view, thereby giving rise to the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.

Simulations with the complete foreground model show the combination of center--heavy features on all antenna spacings from compact emission, and edge--heavy features from both compact and diffuse emission especially the latter even on wide antenna spacings.  This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.

All these observations are consistent with those from Figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}. The observability of the {\it pitchfork} signature predicted in this paper depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise. In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd_data}). We attribute this to differences between our foreground model and the actual sky.  Deeper observations should reveal the feature clearly.

We also note that increasing the antenna spacing results in progressively improving the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}. This improves the localization of foreground objects whose signatures are imprinted in the {\bf delay power spectrum}. For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m. In this case, these correspond to secondary lobes of the power pattern that lie between the primary lobe and the horizon along the local meridian. At short antenna spacings, due to relatively lower resolution along the delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible. These findings are consistent with the discussion in \S\ref{sec:compact}.

In summary, the brightest feature $\sim 10^{12}$~K$^2$(Mpc/$h$)$^3$ is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain. The next brightest signature $\gtrsim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (around $\tau=0$) rather than towards the boundaries. Diffuse emission co--located at the primary lobe of the power pattern is $\sim 10^3$ fainter relative to compact emission from the same region for a $\sim 84$~m antenna spacing. But unlike the latter, diffuse emission decreases rapidly by over four orders of magnitude as the antenna spacing is widened to $\sim 171$~m. However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power pattern as the antenna spacing is increased. This leads to edge--heavy features in the {\bf delay power spectra}. Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the {\bf delay power spectrum} of the foreground sky. We identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.

@@

Here, we will investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination.

>>>>

\begin{figure*}[htb]
\centering
\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith. Antenna shapes used are: dipole (left), phased array (middle), and dish (right). Refer to text for details. The strength and occupancy of the power patterns are correlated with those of delay power spectra. White dotted lines in the delay power spectra mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish. The phased array has intermediate levels of contamination. Foreground power close to the horizon delay limits in all three cases are significant even on long baselines. The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales. The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
\label{fig:aperture-shapes}
\end{figure*}

@@

\begin{figure*}[htb]
\centering
\includegraphics[width=\linewidth]{figure7.eps}
\caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The {\it foreground wedge} is bounded by white dotted lines. Model matches the data to a level consistent with the uncertainties in foreground models and the antenna beam. \label{fig:fhd-sim-comparison}}
\end{figure*}

@@

We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. % Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

@@

\subsection{Analysis of Foreground Signatures}\label{sec:FG-analysis}

@@

\item {\it Sky Model}: Our model of the sky consists of diffuse and compact emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}, respectively. It also consists of localized regions of strong emission such as the Galactic plane and Galactic center. In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission. We will investigate the signatures of each component and their potential impact on {\it EoR window} contamination.

@@

\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the tile power pattern tile changes in any observing mode that tracks the {\bf pointing center}. In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra}.

@@

Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum} shows the delay power spectra obtained from the diffuse (left) and compact (right) foreground emission for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without thermal noise component. Some of the notable signatures are discussed below.

\begin{figure*}[htb]
\centering
\includegraphics[width=\linewidth]{figure8.eps}
\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse (left) and compact (right) foreground models without any thermal noise. The axes and color scale are identical to those in Figure~\ref{fig:fhd-sim-comparison}. In the {\it off-zenith} pointing, emission from the Galactic center is the most prominent feature seen as a branch at $\tau<0$. In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings due to wide--field effects. Compact emission is centrally concentrated. This combination gives rise to a {\it pitchfork} signature in the {\it foreground wedge}. \label{fig:noiseless-dsm-csm-delay-spectrum}}
\end{figure*}

@@

The most prominent signature seen in the {\it off--zenith} pointing (top left panel of Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center is below the horizon.

\subsubsection{Ubiquitous Diffuse Emission}\label{sec:diffuse-features}

Diffuse emission outside the Galactic plane manifests in the primary field of view as a branch at $\tau>0$ and $\tau=0$ in the {\it off--zenith} and {\it zenith} pointings respectively. The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith. As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission. Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.

@@

In both pointings the diffuse emission (left panels) is prominent near the horizon delay limits extending to the widest antenna spacings. This is a characteristic signature of the wide--field effects discussed in \S\ref{sec:wide-field}. It is evident at all LSTs in our simulations. Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines. It decreases in strength with increasing baseline length but is nevertheless present in all baseline orientations.

\subsubsection{Compact Foreground Signatures}\label{sec:compact}

In contrast to the {\bf delay power spectra} of diffuse emission, compact emission (right panels) manifests as a center--heavy structure in either pointing.

The amplitude response of an interferometer to a point source is, to first order, flat cross baseline length. Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary lobe of the power pattern is centered at zenith.

@@

When not dominated by the bright emission from the Galactic center, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, {\bf both of which are significant}. This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines. The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to 
uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figure9.eps}
\caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}). The baseline length and orientation is specified in each panel. The solid red, cyan, and black lines denote contributions from diffuse, compact, and composite foreground models respectively. Vertical dotted lines mark the horizon delay limits. {\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure. The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum. This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside}. \label{fig:pitchfork-baselines}}
\end{figure}

@@

Near the horizon delay limits, the diffuse component is brighter relative to the compact component. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay. In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude. This feature is described in \S\ref{sec:diffuse-long-baselines}. We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.

Simulations with the complete foreground model show the combination of center--heavy features dominated by compact emission in primary field of view, and edge--heavy features from both types of emission especially the diffuse component near the horizon. This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.

The observability of the {\it pitchfork} signature predicted in this paper depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise. In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected in a noisy scenario, although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd-sim-comparison}). We attribute this to differences between our foreground model and the actual sky.  Deeper observations should reveal the feature clearly.

We also note that increasing the antenna spacing progressively improves the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}. This improves the localization of foreground objects whose signatures are imprinted in the {\bf delay power spectrum}. For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between $\tau=0$ and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m. In this case, these correspond to secondary lobes of the power pattern that lie between the primary lobe and the horizon along the local meridian. At short antenna spacings, due to relatively lower resolution along the delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.

In summary, the brightest feature $\sim 10^{12}$~K$^2$(Mpc/$h$)$^3$ is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain. The next brightest signature $\sim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission {\bf from the primary field of view} appearing to be concentrated in the inner regions of the {\it foreground wedge} (around $\tau=0$). Diffuse emission from this region is $\sim 10^3$ fainter relative to that from compact emission for a $\sim 84$~m antenna spacing. But unlike the latter, diffuse emission decreases rapidly by over four orders of magnitude as the antenna spacing is widened to $\sim 171$~m. However, diffuse emission near the horizon is significantly boosted by wide--field effects compared to that in the primary field of view and is prominent across all antenna spacings. This results in a {\it three--pronged pitchfork}--shaped signature inside the {\it foreground wedge} (see also Figure~\ref{fig:antennas-delay-spectra}).

@@

Here, we will investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination. We use the MWA as an example.
